---
title: 1. How has COVID affected the health of the retail industry, as measured by
  employment?
author: "Jill Thompson, Jomaica Alfiler"
date: "8/14/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following analysis attempts to determine how COVID-19 affected the retail industry's health, with retail unemployment used as a measure of industry health. The assumption is that a healthy retail industry has a consistent flow of revenue, allowing it to keep a consistent sized employee pool. An unhealthy industry would be characterized by a high rate of layoffs as retail revenues decline, forcing retailers to minimize losses by reducing employee-related operational costs.


[IPMUS employment data](https://cps.ipums.org/cps/index.shtml) was used in this analysis, with sample dates ranging from January 2019 to January 2021. The following variables were used:

| Variable |  Definition
|:------    |:-------------------------------------------------     | 
| `statefip` | identifies the household's state of residence alphabetically |
| `durunemp` | indicates for how many consecutive weeks each currently unemployed respondent had been without a job and looking for work |
| `empstat` | employment status |
| `whyunemp` | specifies why respondents were unemployed |            |
| `occ2010`  | respondent's occupation based on the Census Bureau's scheme |
| `covidunaw` | could respondent's unemployment be attributed to COVID-19 |
| `cases` | total covid cases |
| `lockdown_start` | stay at home implemented lock down by government |


```{r include=FALSE}
library(tidyverse)
library(jtools)
library(vtable)
library(haven)
library(lubridate)
library(estimatr)
library(fixest)
```

**Raw Data**:
```{r message=FALSE, warning=FALSE}
b_emp <- read_dta('./cps_00001.dta')
b_state_lockdowns <- read_csv('./state_lockdowns.csv')
b_covid_history <- read_csv('./covid_history.csv')
```

```{r include=FALSE}
class_conversion <- function(df, x, y){
  'str_prefix' = x
  'new_class' = y
  df %>% mutate(across(starts_with(str_prefix), new_class))
}
```


```{r}
# data wrangling
t_emp <- b_emp %>% filter(!(whyunemp == 0)) %>%
  filter(occ2010 %in% c(4700:5790)) %>% 
  filter(whyunemp %in% c(1,2,3)) %>% 
  class_conversion(c('month', 'durunemp', 'statefip'), as.numeric) %>% 
  class_conversion(c('occ', 'covidunaw', 'whyunemp'), as.factor) %>% 
  select(-'serial')


t_state_lockdowns <- b_state_lockdowns %>% 
  mutate(lockdown_start = mdy(lockdown_start))

# covid history data
state_name_key <- read.csv('./state_names.csv')

state_name_key <- state_name_key %>% select(-'Abbrev') 

names(state_name_key)[2] <- 'Code' 

names(b_covid_history)[2] <- 'Code' 

t_covid_history <- b_covid_history %>%
  left_join(state_name_key, 'Code')%>% 
  mutate(date = mdy(date)) 


# employment data
exp_emp <- t_emp %>% 
  na.omit()%>%
  mutate(survey_dates = ymd(paste0(year, '-', month, '-01')))%>%
  group_by(year, month, statefip) %>% 
  mutate(layoff = ifelse(whyunemp == 1,1,0), # if reason is layoff = 1, if not = 0
         other_type_loss = ifelse(whyunemp == 2,1,0), 
         temp_emp = ifelse(whyunemp == 3,1,0),
         covidunaw=ifelse(covidunaw==1,1,0))%>% 
  mutate(covid_impact_no = ifelse(covidunaw == 0,1,0), 
         covid_impact_yes = ifelse(covidunaw == 1,1,0))%>% 
  left_join(t_state_lockdowns, 'statefip')


```

### Varibale Analysis

| Variable |  Definition
|:------    |:-------------------------------------------------     | 
| `whyunemp` = "rationale" | this variable is categorized into a "rationale" with reasons being: **"Laid Off"**, **"Other"** and **"Temporary Employment"** which specifies why respondents were unemployed |
| `covidunaw` = "covid_impact" | could respondent's unemployment be attributed to COVID-19. This was categorized into `covid_impact_yes` and `covid_impact_no`. |


Unemployment from our data was categorized into three "rationale": 
Laid Off, Temporary Employment, and Other.
In which we wanted to look at whether there was an increase of unemployment 
in one category over the others. 


From this graph, our results shows that there was an increase in unemployment 
right around when COVID-19 cases came to the US - which was in March, with a peak in April. 
We also identified that the majority of unemployment came from workers who were "Laid Off". 
We also see minimal impact of unemployment for "Temporary workers". 
Finally, at the end of 2020, we see that those who are unemployed for "Other" reasons had an increase 
and becomes the majority reason over being laid off. 

```{r}
sum_by_why <- t_emp %>%
  mutate(date = ymd(paste0(year, '-', month, '-01'))) %>% 
  dplyr::count(date, whyunemp) %>%
  filter(date >= "2019-12-01")

plot_by_why <- sum_by_why %>%
  ggplot(aes(x = date, y = n, fill = factor(whyunemp))) +
  geom_bar(position = "dodge", stat = "identity") +
  theme(legend.position="bottom") + 
  xlab("Month") +
  ylab("Number of People") +
  ggtitle('Figure 1. Types of Unemployment by Month')+
  scale_fill_manual(values=c("#CC6666", "#9999CC", "#66CC99"), name = "Reason for Unemployment: ",
                      labels=c("Laid Off", "Other", "Temporary Employment"))
plot_by_why
```

We also plotted the total number of unemployment cases reported and if the reason was COVID related. From this graph, we can see that the majority of unemployment reported in this time frame is due to COVID.

From this graph, we can see that COVID was the majority reason for why there was unemployment. 
The "Not Reason For" category does not appear to have been effected much when looking at data in 
the last two months of 2019. 

```{r}
sum_unemp <- t_emp %>%
  na_if(99) %>%
  mutate(date = ymd(paste0(year, '-', month, '-01'))) %>% 
  dplyr::count(date, covidunaw) %>%
  filter(covidunaw %in% c(NA, 1, 2)) %>%
  filter(date >= "2019-11-01") %>% 
  mutate_all(funs(ifelse(is.na(.), 0, .)))

plot_total <- ggplot(data = sum_unemp, aes(x = date, y = n, fill = factor(covidunaw))) +
  geom_bar(position = "stack", stat = "identity") +
  scale_fill_manual(name = "Covid Unemployment: ",
                      labels = c("Not Recorded", "Not Related", "COVID Related"),
                      values = c('gray', "steel blue", "red")) +
  theme(legend.position="bottom") +
  ggtitle('Figure 2. Covid-19 Attributed Unemployment by Month')
plot_total
```

Seeing the spike in unemployment at the beginning of 2020 in Figure 1. above and Figure 2., we established an initial hypothesis that COVID-19 measures imposed to reduce the spread of the disease were negatively impacting the retail industry’s health. Primarily, we looked at lockdown orders:

**H0** = As measured by unemployment, COVID had no impact on retail industry health. (*Key*: COVID had no impact.)

**H1** = Retail industry health was affected by COVID related measures, specifically stay at home orders. (*Key*: COVID related measures: lock downs.)


The following regression model assessed if there was a relationship between Lockdown start dates, covid related unemployment and retail layoffs. The model included a LSDV factor for states to provide a fixed effect control for collective variables such as politics, which could have influenced COVID-19 government policy implementation during that time.
```{r}
# Regression

# model of state lockdown order start dates, impact of covid on unemployment, state and layoffs.
m1_state_closure_layoff <- lm_robust(layoff ~ lockdown_start + covid_impact_yes + factor(state), data = exp_emp)

# Find any relationship from model 1.
export_summs(m1_state_closure_layoff, model.names = 'Covid Related Unemployment v.\nLayoffs v. Lockdown')
```
In the first model, you can see that lock down orders had a significant relationship between layoffs, lockdowns and COVID related unemployment due to the states being factored. This is because some states remained open, while some didn't which could explain why there was high rates of unemployment due to the states being closed down. 
To better understand what may have led to this relationship, we have decided to factor out the states. 
For this analysis to occur we are assuming that lock down orders did not have an effect on unemployment. 

```{r}
# 2nd Regression.

# model of state lockdown order start dates, impact of covid on unemployment, and layoffs.
m2_state_closure_layoff <- lm_robust(layoff ~ lockdown_start + covid_impact_yes, data = exp_emp)

# Find any relationship from model 1.
export_summs(m2_state_closure_layoff, model.names = 'Covid Related Unemployment v.\nLayoffs v. Lockdown')
```
The model did not show a statistically significant relationship between the lock down start dates, layoffs and COVID related unemployment. This is hardly surprising given that layoffs began to reduce shortly after the first government lock down orders were issued on March 18th, 2020, as seen by the plot. The IPMUS variable "covid_impact" exhibits this tendency as well (see above). Beginning in April 2020, IPUMS began asking respondents if their unemployment was due to COVID. This statistic also fell, mirroring the drop in unemployment caused by layoffs following the first lock down. Furthermore, lockdown orders does not appear to have had a negative impact on the retail sector. Thus, we fail to reject our null and cannot accept the alternative hypothesis.

---
## Additional analysis
Next, we decided to look at the data using unemployment rates within the data as our metric (rather than number of layoffs).

```{r message=FALSE, warning=FALSE}
employment_data <- read_dta('cps_00010.dta')

states <- read_csv('state_codes.csv')
```

Filter for only desired employment variables and values:

  `empstat`:  Values between 10 and 29 relate to people in the labor market.
  `ind`:  Values between 4670 and 5790 indicate the retail industry
  
```{r message=FALSE, warning=FALSE}
t_employment <- employment_data %>% 
  filter(ind>=4670 & ind<=5790) %>% 
  filter(empstat < 30 & empstat >= 10) %>% 
  mutate(emp_Status = empstat, industry = ind)
```

Create a binary variable to indicate employment status, 0 = unemployed,
 1 = employed

```{r message=FALSE, warning=FALSE}
t_employment <- t_employment %>% 
  mutate(employed = ifelse(empstat < 20, 1, 0))
```

Create a variable to indicate whether data point is from before or after
 COVID, 0 = pre-covid, 1 = post-covid
Add a year_month variable for grouping during analysis

```{r message=FALSE, warning=FALSE}
t_employment <- t_employment %>% 
  mutate(month = as.factor(month)) %>% 
  mutate(year_month = ym(paste0(year, '-', month))) %>% 
  mutate(post_covid = ifelse(year_month > '2020-03-01', 1, 0))
```

Select only desired variables.

```{r message=FALSE, warning=FALSE}
t_employment <- t_employment %>% 
  select(year, month, employed, year_month, post_covid, statefip) %>% 
  mutate(statefip = as.numeric(statefip)) %>% 
  mutate(for_count = 1)
```

Replace numeric state codes with state names.

```{r message=FALSE, warning=FALSE}
t_employment <- t_employment %>% 
  left_join(states, "statefip") %>% 
  mutate(state = statename) %>% 
  select(-statefip, -statename)
```

Group and summarise data by date only.

```{r message=FALSE, warning=FALSE}
t_employment_group <- t_employment %>% 
  group_by(year_month) %>% 
  summarise(num_employable = sum(for_count),
            num_employed = sum(employed==1),
            num_unemployed = num_employable - num_employed) %>% 
  mutate(percent_unemployed = num_unemployed/num_employable,
         post_covid = ifelse(year_month > '2020-03-01', 1, 0))
```

Group and summarise by date and state. We want to look at data
 grouped by state because we suspect that a lot back-door variation
 will come from economic and political differences between states.

```{r message=FALSE, warning=FALSE}
t_employment_group_2 <- t_employment %>% 
  group_by(state, year_month) %>% 
  summarise(num_employable = sum(for_count),
            num_employed = sum(employed==1),
            num_unemployed = num_employable - num_employed) %>% 
  mutate(percent_unemployed = num_unemployed/num_employable,
         post_covid = ifelse(year_month > '2020-03-01', 1, 0))
```

Before running a regression, we'll graph unemployment trends across time, both
 grouped by state and not grouped by state.

The first graph shows a clear jump in unemployment in April, right
 after COVID. Unemployment levels decrease steadily after that, returning to
 near normal until at the end of 2021.

```{r message=FALSE, warning=FALSE}
t_employment_group %>% 
  ggplot(aes(x=year_month, y=percent_unemployed)) + 
  geom_point()
```

If we run the same graph but with data points grouped by state, we
 see a similar trend overall, but a significant amount of variation
 each month based on state. This is an indicator that we should
 control for state in our regression.

```{r message=FALSE, warning=FALSE}
t_employment_group_2 %>% 
  ggplot(aes(x=year_month, y=percent_unemployed, group=state)) + 
  geom_point
```

Given that we have panel data and our graphs indicate a lot of
 between-state variation, we chose a fixed effects model to look at
 pre- and post-covid changes while controlling for state variations.

```{r message=FALSE, warning=FALSE}
m1 <- t_employment_group_2 %>% 
  feols(percent_unemployed~post_covid | state)

etable(m1)
```

Looking at our results, we can conclude that, when controlling for state,
 monthly average post-covid unemployment rates are 4.67 percent
 higher than average monthly pre-covid unemployment rates. This
 is true at a .001 level of significance.

## Assumption Acknowledgement

1. Our first assumption was that we can control for many economic and political differences by controlling for state.

2. Based on the reports during this timeframe, we assume that the total COVID cases reported are accurate. While higher cases of COVID may have occurred earlier in 2020, they may have gone unreported. We are assuming that all data for COVID cases is correct.

3. We assume unemployment is the main indicator of retail industry performance. It is assumed that a healthy retail industry has a steady flow of customers contributing to steady revenue flow, enabling it to maintain uniformly sized employee pools.


## Final Results: “How Did COVID Impact the Retail Industry as Measured by Unemployment?”

Our first regression shows a statistically significant relationship between layoffs, lockdown start-dates, and COVID-related unemployment.

Our final regression shows a statistically significant increase in monthly average unemployment rates post-Covid (April 2020 - December 2020) versus pre-Covid (January 2019 - March 2020).

We concluded that a rise in unemployment in 2020 was a direct result of COVID-19's negative impact on the retail industry's health. In addition, unemployment was rising long before states started requiring lock down orders. Finally, we observe that before positive cases started to surge upward, unemployment climbed and then fell back to a much lower rate.





