---
title: 1. How has COVID affected the health of the retail industry, as measured by
  employment?
author: "Jomaica Alfiler"
date: "8/14/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following analysis attempts to determine how COVID-19 affected the retail industry's health and retail unemployment is used as a measure of industry health in this analysis. The assumption is that a healthy retail industry has a consistent flow of revenue, allowing it to keep a consistent sized employee pool. An unhealthy industry would be characterized by a high rate of layoffs as retail revenues decline, forcing retailers to minimize losses by reducing employee-related operational costs.


### How Did COVID Impact Unemployment? 

[IPMUS employment data](https://cps.ipums.org/cps/index.shtml) was used in this analysis, with sample dates ranging from January 2019 to January 2021. The following variables were used:

| Variable |  Definition
|:------    |:-------------------------------------------------     | 
| `statefip` | identifies the household's state of residence alphabetically |
| `durunemp` | indicates for how many consecutive weeks each currently unemployed respondent had been without a job and looking for work |
| `empstat` | employment status |
| `whyunemp` | specifies why respondents were unemployed |            |
| `occ2010`  | respondent's occupation based on the Census Bureau's scheme |
| `covidunaw` | could respondent's unemployment be attributed to COVID-19 |
| `lockdown_start` | stay at home implemented lock down by government |


```{r include=FALSE}
library(tidyverse)
library(jtools)
library(vtable)
library(haven)
library(lubridate)
library(estimatr)
```

**Raw Data**:
```{r message=FALSE, warning=FALSE}
q1_employment <- read_dta('cps_00001.dta') 
covid_history <- read_csv('covid_history.csv') 
state_lockdowns <- read_csv('state_lockdowns.csv')
state_names <- read_csv('state_names.csv')
timeline <- read_csv('timeline.csv')
```

**Analysis**:

| Variable |  Definition
|:------    |:-------------------------------------------------     | 
| `whyunemp` = "rationale" | this variable is categorized into a "rationale" with reasons being: **"Laid Off"**, **"Other"** and **"Temporary Employment"** which specifies why respondents were unemployed |
| `occ2010` = "occupation" | respondent's occupation based on the Census Bureau's scheme |
| `covidunaw` = "covid_impact" | could respondent's unemployment be attributed to COVID-19 |

**H0** = As measured by unemployment, COVID had no impact on retail industry health. (*Key*: COVID had no impact.)
**H1** = Retail industry health was affected by COVID related measures, specifically stay at home orders. (*Key*: COVID related measures: lock downs.)


to-do:

* first plot a graph of the total reported unemployment per month:

  + narrow the date range from November 2020 to January 2021, focusing on the time when COVID-19 became a prominent news item.
  
  
* categorize the `whyunemp` data into "rationale" (for reasons why they are unemployed) categories to look at which employment changed into three levels:

  + “Laid Off”,
  + “Other”,
  + “Temporary Job Ended”
  
  
* relationship between `timeline` (google search trends) and layoffs.

* relationship between `lockdown_states` (stay at home order by govt.) and unemployment.

* implement H0 and HA.




