---
title: "DTC Q3"
author: "Erik Campbell, Kristen Higashi"
date: "8/22/2022"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
editor_options: 
  chunk_output_type: inline
---

  
```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = TRUE)

```

# Purpose Statement
The purpose of this document is to investigate the effect of COVID on working individuals across industries and regions. Retail needs to worry about who has money to spend. In order to have money, you need to make money. Thus our focus is to determine who is working, what they are doing, and how that activity has changed across time. 

# Dataset Overview
Our analysis applies 3 main data sources:


1.  IPUMS variables
 IPUMS employment data ([link](https://cps.ipums.org/cps/index.shtml)) - Data collected between January 2019 and Januaray 2021
 
| Variable |  Definition
|:------    |:-------------------------------------------------     | 
| `YEAR` | Year survey was taken |
| `MONTH` | Month survey was taken |
| `EMPSTAT` | Employment status |
| `WTFINL` | Survey person weight (as calculated by CPS) |
| `IND`  | Respondent's work industry code |
| `REGION` | Respondents location (regional) |
| `AGE`  | Respondent's age |
| `SEX` |  Marker for respondent sex assigned at birth |
| `RACE` | Marker for respondent's race |

2.  Provided industry csv (*indnames.csv*)
  Converts IPUMS industry codes to industry names 
 
| Variable |  Definition
|:------    |:-------------------------------------------------     | 
| `IND` | Work industry code (joins to IPUMS IND field) |
| `indname` | Work industry English name |


3.  Provided geographic csv (*state-geocodes-v2021.csv*)
  Converts IPUMS geographies (STATEFIP) to readable geo names.

| Variable |  Definition
|:------    |:-------------------------------------------------     | 
| `STATEFIP` | State FIP code (joins to IPUMS STATEFIP field) |
| `State` | State English name |


The below is in relation to the unemployment rate in Agriculture.
| Industry |  % Change
|:------    |:-------------------------------------------------     | 
|`Arts,Entertainment,andRecreation,andAccommodationandFoodServices`                  | -%05.22 |
|`Construction`                                                                      | -%00.95 |
|`EducationalServices,andHealthCareandSocialAssistance`                            | %01.59 |
|`FinanceandInsurance,andRealEstateandRentalandLeasing`                            | %02.37 |
|`Information`                                                                     | %00.02 |
|`Manufacturing`                                                                  | %00.81 |
|`OtherServices,ExceptPublicAdministration`                                          | -%00.10 |
|`Professional,Scientific,andManagement,andAdministrativeandWasteManagementServices` | %00.32 |
|`PublicAdministration`                                                            | %03.07 |
|`RetailTrade`                                                                 | -%00.66 |
|`TransportationandWarehousing,andUtilities`                                         | -%00.13 |
|`WholesaleTrade`                                                                  | %01.36 |
|`period_typeCovid`                                                                  | -%00.81 |
|`period_typePost_Covid`                                                            | -%2.92


## Load Libraries
```{r, warning = FALSE, message = FALSE}
# Load Relevent Libraries
library(ipumsr)  
library(dplyr)
library(tidyverse)
library(jtools)
library(vtable)
library(lubridate)
library(haven)
library(fixest)
library(ggplot2)
library(formattable)
library(echarts4r)
library(ggplot2)
```

## Load and Mutate Data

```{r, warning = FALSE, message = FALSE}
#### READ IN INDUSTRY CODES ####
Industries <- read_csv('indnames.csv') %>%
  rename(IND = ind)
###### Read in State Federal Information Processing Codes. ######  
state_fips <- read_csv('state-geocodes-v2021.csv') %>%
  filter(!is.na(Region)) %>%
  rename(State = Name)
  #select(STATEFIP, State)
###### read in IPUMS demographic data from
dem_ddi <- read_ipums_ddi("cps_00016.xml")
dem_data <- read_ipums_micro(dem_ddi) %>% 
  zap_labels() %>%  
  mutate(
    month_digit= case_when( 
      nchar(str_trim(as.character(MONTH))) ==1 ~ (paste('0',str_trim(as.character(MONTH)), sep = '')),
      nchar(str_trim(as.character(MONTH))) ==2 ~ (str_trim(as.character(MONTH))), 
      TRUE ~ 'WRONG'
    )
    , date_ = as_date(paste(YEAR, month_digit, '01', sep = '-'))
    , period_type = as_factor(case_when(
      date_ < '2020-03-01' ~'Pre_Covid', 
      date_ >= '2020-03-01' & date_ < '2020-04-01' ~'Covid', 
      date_ >= '2020-04-01' ~'Post_Covid'
    ))
  )
#### JOIN FILES INTO ONE ####
DF <- dem_data %>% 
  ## Join in geographic info
  left_join(state_fips, by = 'STATEFIP') %>% # 3.9 million observations
  ## Join in industry categories
  left_join(Industries, by = 'IND')
#### DEMOGRAPHICS DATAFRAME ####
demo_data <- DF %>%
  mutate(EmploymentStatus = case_when(
    EMPSTAT <= 1 | EMPSTAT >= 30 ~  'Other',
    EMPSTAT >= 10 & EMPSTAT < 20 ~ '1',
    EMPSTAT >= 20 & EMPSTAT < 30 ~ '0')) %>%
  mutate(sex = case_when(
    SEX == 1 ~  'male',
    SEX == 2 ~ 'female')) %>%
  mutate(MARST = case_when(
    MARST == 1 ~ 'Married',
    MARST != 1 ~ 'Single')) %>%
  mutate(RACE = case_when(
    HISPAN == '0' & RACE == '100' ~ 'White',
    HISPAN == '0' & RACE == '200' ~ 'Black',  
    HISPAN == '0' & RACE == '300' ~ 'AmericanIndian',
    HISPAN == '0' & RACE %in% c('650', '651', '652') ~ 'Asian',
    HISPAN == '0' & !(RACE %in% c('100', '200', '300', '650', '651', '652')) ~ 'MixedRace',
    HISPAN != '0' ~ 'Latino')) %>%
  mutate(AgeGroup = case_when(
    AGE < 20 ~ 'under20',
    AGE >= 20 & AGE < 30 ~ '20to29',
    AGE >= 30 & AGE < 40 ~ '30to39',
    AGE >= 40 & AGE < 50 ~ '40to49',
    AGE >= 50 & AGE < 60 ~ '50to59',
    AGE >= 60 & AGE <= 65 ~ '60to65',
    AGE >= 65 ~ 'RetirementAge')) %>%
  mutate(REGION = case_when(
    REGION == 11 | REGION == 12 ~ 'Northeast',
    REGION == 21 | REGION == 22 ~ 'Midwest',
    REGION == 31 | REGION == 32 |  REGION == 33 ~ 'South',
    REGION == 41 | REGION == 42 ~ 'West',
    REGION == 97 ~ 'Unknown')) %>%
  mutate(METRO = case_when (
    METRO %in% c(0, 4) ~  'Other',
    METRO == 1 ~  'Rural',
    METRO == 2 ~  'CityCenter',
    METRO == 3 ~  'Suburb')) %>%
  mutate(Education = case_when(
    EDUC99 == 00 ~ 'Unknown',
    EDUC99 >= 1 & EDUC99 <=9 ~ 'NoDiploma',
    EDUC99 == 10 ~ 'HighSchoolDiploma',
    EDUC99 == 11 ~ 'SomeCollege',
    EDUC99 >= 12 & EDUC99 <= 14 ~  'AssociateDegree',
    EDUC99 == 15 ~ 'Bachelors',
    EDUC99 == 16 ~ 'Masters',
    EDUC99 > 16 ~ 'PhDProfessional')) %>%
  mutate(LaborForce = case_when(
    LABFORCE == 0  ~ 'Military',
    LABFORCE == 1  ~ '0',
    LABFORCE == 2  ~ '1')) %>%
  mutate(WorkStatus = case_when(
    WKSTAT %in% c(11, 14, 15) ~ 'FullTime',
    WKSTAT %in% c(12, 20, 21, 22, 40, 41) ~ 'PartTime',
    WKSTAT == 13 ~ 'FullTimeNoWork',
    WKSTAT == 42 ~ 'PartTimeNoWork',
    WKSTAT == 50 ~ 'NoWorkSeekingFullTime',
    WKSTAT == 60 ~ 'NoWorkSeekingPartTime',
    WKSTAT == 99 ~ 'MilitaryNoLaborForce',)) %>%
  mutate(FamilyIncome =case_when(                        
    FAMINC <= 600 ~ '<$25k_FPL_Familyof4',
    FAMINC > 600 & FAMINC <= 740 ~ '25k-50K',
    FAMINC > 740 & FAMINC <= 830 ~ '50k-75k',
    FAMINC > 830 & FAMINC <= 841~ '75k-100k',
    FAMINC == 842  ~ '100k - 150k',
    FAMINC == 843 ~ '150k+',
    FAMINC == 999 ~ 'NA')) %>%
  mutate(WHYUNEMP = case_when(
    WHYUNEMP <= 3 ~ 'LostJob',
    WHYUNEMP == 4 ~ 'Quit',
    WHYUNEMP == 5 ~ 'ReEnterLaborForce',
    WHYUNEMP == 6 ~ 'FirstJob',
    WHYUNEMP == 0 ~ 'Military')) %>%
  mutate(employed = ifelse(EMPSTAT >= 10 & EMPSTAT <20, 1, 0), unemployed = ifelse(EMPSTAT >= 20 & EMPSTAT<= 22, 1, 0)) %>% # In here twice!!!, I updated the conditions to match IPUMS EMPSTAT definitions. 
  mutate(is_employed = ifelse(EMPSTAT>=10 & EMPSTAT<20, TRUE, FALSE)) %>%  
  mutate(retail_v_other = ifelse(indname=='Retail Trade', 'Retail', 'Other') ## flag if industry == Retail
  )%>% 
  filter(EMPSTAT < 30 & EMPSTAT >= 10) %>% 
  mutate_if(is.character, as.factor) %>%
  unite(HHID, HRHHID, HRHHID2) %>%
  drop_na(indname) %>%  # Dropping N/A's in indname as NIU's in military.
  select(date_, YEAR, MONTH, HHID, PERNUM, EMPSTAT, is_employed, employed, unemployed, CPSID, CPSIDP
          , HWTFINL, WTFINL, FAMINC, REGION, METRO, AGE, AgeGroup, SEX, sex
          , RACE, MARST, IND, WHYUNEMP, AgeGroup, Education, LaborForce
          , EmploymentStatus, WorkStatus, FamilyIncome, indname, retail_v_other, State, period_type)
```


## EDA

```{r, warning = FALSE, message = FALSE}
# distinct people in survey = 500,130
summary(demo_data)
vtable(demo_data, lush = TRUE)
reg_df_vis<- demo_data %>% 
  select(period_type
         , date_
         ## Measure
         , is_employed
         , WTFINL
         ## Variables
         , IND, indname, REGION
         ## Fixed
         , AGE, sex, RACE
  )
```

```{r, warning = FALSE, message = FALSE}
## All industries unemployment rate over time
unemp_time <- demo_data %>% 
  filter(indname != 'Military', indname != 'Other') %>%
  group_by(date_) %>%
  summarize(UnemploymentRate = 1- (weighted.mean(is_employed == TRUE, w = WTFINL))) %>% 
  ggplot(aes(x = date_, y = UnemploymentRate)) + geom_line()+
  geom_rect(aes(xmin=as_date('2020-03-01'), xmax=as_date('2020-04-01'), ymin=-Inf, ymax=Inf),  linetype = 'dashed',color="light green", fill = 'light green' ,alpha=.01)+
  scale_y_continuous(labels = scales::percent_format(accuracy = .1))+
  labs(x = "Survey Date", y = "Unemployment Rate",
    title ="Unemployment By Industry")+
  theme(aspect.ratio=3/7)
unemp_time

## All industries v retail unemployment rate over time
unemp_retail_time <- demo_data %>%
  filter(indname != 'Military', indname != 'Other') %>% 
  group_by(date_, retail_v_other) %>%
  summarize(UnemploymentRate = 1- (weighted.mean(is_employed == TRUE, w = WTFINL))) %>% 
  ggplot(aes(x = date_, y = UnemploymentRate, color = retail_v_other)) + geom_line()+
  geom_rect(aes(xmin=as_date('2020-03-01'), xmax=as_date('2020-04-01'), ymin=-Inf, ymax=Inf),  linetype = 'dashed',color="light green", fill = 'light green' ,alpha=.01)+
  labs(x = "Survey Date", y = "Unemployment Rate",
    title ="Unemployment (Retail v Other Industries)")+
  scale_y_continuous(labels = scales::percent_format(accuracy = .1))+
  theme(aspect.ratio=3/7)
unemp_retail_time
## limit x window for clarity
unemp_time_skinny <- demo_data %>% 
  filter(date_>=as_date('2020-01-01'),
         date_<= as_date('2020-06-01')) %>%
  filter(indname != 'Military', indname != 'Other') %>%
  group_by(date_, retail_v_other) %>%
  summarize(UnemploymentRate = 1- (weighted.mean(is_employed == TRUE, w = WTFINL))) %>% 
  ggplot(aes(x = date_, y = UnemploymentRate, color = retail_v_other)) + geom_line()+
  geom_rect(aes(xmin=as_date('2020-03-01'), xmax=as_date('2020-04-01'), ymin=-Inf, ymax=Inf),  linetype = 'dashed',color="light green", fill = 'light green' ,alpha=.01)+
  labs(x = "Survey Date", y = "Unemployment Rate",
    title ="Unemployment (Retail v Other Industries), 2020")+
  scale_y_continuous(labels = scales::percent_format(accuracy = .1))+
  theme(aspect.ratio=3/7)
unemp_time_skinny
```

```{r, warning = FALSE, message = FALSE}
unemp_industries <- demo_data %>%
  filter(indname != 'Military', indname != 'Other') %>%
  group_by(date_, indname) %>%
  summarize(UnemploymentRate = 1- (weighted.mean(is_employed == TRUE, w = WTFINL))) %>% 
  ggplot(aes(x = date_, y = UnemploymentRate, color = indname)) + geom_line()+
  geom_rect(aes(xmin=as_date('2020-03-01'), xmax=as_date('2020-04-01'), ymin=-Inf, ymax=Inf),  linetype = 'dashed',color="light green", fill = 'light green' ,alpha=.01)+ labs(x = "Survey Date", y = "Unemployment Rate",
    title ="Unemployment All Industries")+
  scale_y_continuous(labels = scales::percent_format(accuracy = .1))+
  theme(aspect.ratio=3/7, legend.position = 'none')
unemp_industries
```

### Geographical
```{r, warning = FALSE, message = FALSE}
unemp_by_region <- demo_data %>%
  filter(indname != 'Military', indname != 'Other') %>% 
  group_by(date_, REGION) %>%
  summarize(UnemploymentRate = 1- (weighted.mean(is_employed == TRUE, w = WTFINL))) %>% 
  ggplot(aes(x = date_, y = UnemploymentRate, color = REGION)) + geom_line()+
  geom_rect(aes(xmin=as_date('2020-03-01'), xmax=as_date('2020-04-01'), ymin=-Inf, ymax=Inf),  linetype = 'dashed',color="light green", fill = 'light green' ,alpha=.01)+
  labs(x = "Survey Date", y = "Unemployment Rate",
    title ="Unemployment by Region All Industries")+
  scale_y_continuous(labels = scales::percent_format(accuracy = .1))+
  theme(aspect.ratio=3/7)
unemp_by_region

unemp_by_metro <- demo_data %>%
  filter(indname != 'Military', indname != 'Other') %>% 
  group_by(date_, METRO) %>%
  summarize(UnemploymentRate = 1- (weighted.mean(is_employed == TRUE, w = WTFINL))) %>% 
  ggplot(aes(x = date_, y = UnemploymentRate, color = METRO)) + geom_line()+
  theme(aspect.ratio=3/7)+
  geom_rect(aes(xmin=as_date('2020-03-01'), xmax=as_date('2020-04-01'), ymin=-Inf, ymax=Inf),  linetype = 'dashed',color="light green", fill = 'light green' ,alpha=.01)+
  labs(x = "Survey Date", y = "Unemployment Rate",
    title ="Unemployment by Location : City, Suburban, Rural, Other")+
  scale_y_continuous(labels = scales::percent_format(accuracy = .1))+
  theme(aspect.ratio=3/7)
unemp_by_metro
```

### Race
```{r, warning = FALSE, message = FALSE}
#### RACE PRE-COVID EMPLOYMENT ####
unemp_by_race <- demo_data %>%
  filter(indname != 'Military', indname != 'Other') %>% 
  group_by(date_, RACE) %>%
  summarize(UnemploymentRate = 1- (weighted.mean(is_employed == TRUE, w = WTFINL))) %>% 
  ggplot(aes(x = date_, y = UnemploymentRate, color = RACE)) + geom_line()+
  theme(aspect.ratio=3/7)+
  geom_rect(aes(xmin=as_date('2020-03-01'), xmax=as_date('2020-04-01'), ymin=-Inf, ymax=Inf),  linetype = 'dashed',color="light green", fill = 'light green' ,alpha=.01)+
  labs(x = "Survey Date", y = "Unemployment Rate",
    title ="Unemployment All Industries")+
  scale_y_continuous(labels = scales::percent_format(accuracy = .1))+
  theme(aspect.ratio=3/7)
unemp_by_race
```

### Age
```{r, warning = FALSE, message = FALSE}
unemp_by_age_group <- demo_data %>%
  filter(indname != 'Military', indname != 'Other') %>% 
  group_by(date_, AgeGroup) %>%
  summarize(UnemploymentRate = 1- (weighted.mean(is_employed == TRUE, w = WTFINL))) %>% 
  ggplot(aes(x = date_, y = UnemploymentRate, color = AgeGroup)) + geom_line()+
  theme(aspect.ratio=3/7)+
  geom_rect(aes(xmin=as_date('2020-03-01'), xmax=as_date('2020-04-01'), ymin=-Inf, ymax=Inf),  linetype = 'dashed',color="light green", fill = 'light green' ,alpha=.01)+
  labs(x = "Survey Date", y = "Unemployment Rate",
    title ="Unemployment All Industries")+
  scale_y_continuous(labels = scales::percent_format(accuracy = .1))+
  theme(aspect.ratio=3/7)
unemp_by_age_group
```

# Regressions
```{r, warning = FALSE, message = FALSE}
reg_df <- demo_data %>%
  select(period_type
         , date_
         ## Measure
         , is_employed
         , WTFINL
         ## Variables
         , IND, indname, REGION
         ## Fixed
         , AGE, sex, RACE, AgeGroup, MARST
  )

reg1 <- feols(is_employed~period_type + indname, se = 'hetero', data = reg_df)
reg1a <- feols(is_employed~period_type, se = 'hetero', data = reg_df)
etable(reg1, reg1a)               
```
--Interpretation reg1a =  
-controlling for industry and relative to agricultural industry, a one-unit change in pre-covid employment is associated with a 2.92% decrease within the Entertainment and Recreation industry.

-controlling for industry and relative to agricultural industry, a one-unit change in pre-covid employment is associated with a 5.22% decrease in within the Entertainment and Recreation industry.

-controlling for industry and relative to agricultural industry, a one-unit change in pre-covid employment is associated with a .95% decrease within the Educational Services industry.

-controlling for industry and relative to agricultural industry, a one-unit change in pre-covid employment is associated with a 1.59% increase within the Educational Services industry.

-controlling for industry and relative to agricultural industry, a one-unit change in pre-covid employment is associated with a 2.37% increase within the Financial & Insurance services industry.

-controlling for industry and relative to agricultural industry, a one-unit change in pre-covid employment is associated with a .02% increase within the Information services industry.

-controlling for industry and relative to agricultural industry, a one-unit change in pre-covid employment is associated with a .81% increase within the Manufacturing industry.

-controlling for industry and relative to agricultural industry, a one-unit change in pre-covid employment is associated with a -95.02% decrease within the Manufacturing industry.  # logically work through this, does it make sense?  Simply filter out?

-controlling for industry and relative to agricultural industry, a one-unit change in pre-covid employment is associated with a .1% decrease within Other Services Industry.

-controlling for industry and relative to agricultural industry, a one-unit change in pre-covid employment is associated with a .32% increase within Professional, Scientific and Management Industry.

-controlling for industry and relative to agricultural industry, a one-unit change in pre-covid employment is associated with a 3.7% increase within Public Administration Industry.

-controlling for industry and relative to agricultural industry, a one-unit change in pre-covid employment is associated with a .66% decrease within Retail Trade Industry.

-controlling for industry and relative to agricultural industry, a one-unit change in pre-covid employment is associated with a .13% decrease within Transportation & warehousing Industry.

-controlling for industry and relative to agricultural industry, a one-unit change in pre-covid employment is associated with a 1.36% increase within Wholesale Trade Industry.

'''
reg2 <- feols(is_employed~indname + period_type, se = 'hetero', data = reg_df)
etable(reg2)

```
Interpretation reg2- note:summary table above in findings.
-controlling for period_tyep[pre-covid(1/2019-2/2020), covid(3/2020), and post covid(4/2020-12/2021)] and relative to agricultural industry, a one-unit change in industry employment is associated with a -5.22% decrease within the Entertainment and Recreation industry.
 
-controlling for period_tyep[pre-covid(1/2019-2/2020), covid(3/2020), and post covid(4/2020-12/2021)] and relative to agricultural industry, a one-unit change in industry employment is associated with a -.66% decrease within the Retail Trade industry.
 
-controlling for period_tyep[pre-covid(1/2019-2/2020), covid(3/2020), and post covid(4/2020-12/2021)] and relative to agricultural industry, a one-unit change in industry employment is associated with a -.13% decrease within the Transportation & warehousing industry.

# Assumptions
We assumed that the number of jobs in an industry and money a family has to spend on items are equivalent within all industry categories.  This is not necessarily true as not all jobs are created equal in terms of amount of discretionary money a family has to spend on goods and services.  We've not taken into account household savings and debt rates from Jan 2019 through Dec 2021, and these factors play a role in discretionary income families have to spend.  In the future, we'd take into account and create faminly income on the monthly level and not just industry members of household worked in, but account for their job.  In retrospect, accounting for jobs within industries that can be performed from an employee's home would assist in identifying and quantifying how many earners moved from cities to suburbs or rural areas.  Cohort that relocated during the pandemic to suburbs, working in high-earninging occupations, present an opportunity for businesses to meet customers demands at different geographic locations.  