---
title: "DTC Q3"
author: "Erik Campbell, Kristen Higashi"
date: "8/22/2022"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
editor_options: 
  chunk_output_type: inline
---

  
```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = TRUE)

```

# Purpose Statement
The purpose of this document is to investigate the effect of COVID on working individuals across industries and regions. Retail needs to worry about who has money to spend. In order to have money, you need to make money. Thus our focus is to determine who is working, what they are doing, and how that activity has changed across time. 

# Dataset Overview
Our analysis applies 3 main data sources:


1.  IPUMS variables
 IPUMS employment data ([link](https://cps.ipums.org/cps/index.shtml)) - Data collected between January 2019 and Januaray 2021
 
| Variable |  Definition
|:------    |:-------------------------------------------------     | 
| `YEAR` | Year survey was taken |
| `MONTH` | Month survey was taken |
| `EMPSTAT` | Employment status |
| `final_weight` | Survey person weight (as calculated by CPS) |
| `IND`  | Respondent's work industry code |
| `REGION` | Respondents location (regional) |
| `AGE`  | Respondent's age |
| `SEX` |  Marker for respondent sex assigned at birth |
| `RACE` | Marker for respondent's race |

2.  Provided industry csv (*indnames.csv*)
  Converts IPUMS industry codes to industry names 
 
| Variable |  Definition
|:------    |:-------------------------------------------------     | 
| `IND` | Work industry code (joins to IPUMS IND field) |
| `indname` | Work industry English name |


3.  Provided geographic csv (*state-geocodes-v2021.csv*)
  Converts IPUMS geographies (STATEFIP) to readable geo names.

| Variable |  Definition
|:------    |:-------------------------------------------------     | 
| `STATEFIP` | State FIP code (joins to IPUMS STATEFIP field) |
| `State` | State English name |



# Background & Assumptions
Our study was shaped using unemployment figures by industry, geography, race and age from January 2019 through December 2021.  Operating under the guidance that anything before March 2020 was "Pre-Covid," March 2020 was "Covid", and anything after March 2020 was "Post-Covid," we broke our data into three distinct time periods.  These time periods were Pre-Covid (JAN 2019-through Feb 2020), Covid (March 2020) and Post-Covid (April 2020-through- December 2021) and evaluated changes in unemployment over time by industry, changes to unemployment by geography, race and age.  We made assumptions when evaluating industry unemployment that jobs numbers in relation to different industries are similar; however, employment in one industry(e.g. technology, healthcare) versus employment in the leisure and hospitality industry lead to differing amounts of discretionary income available to consumers.

We did not take into account household savings and debt rates from Jan 2019 through Dec 2021, and these factors play a role in discretionary income families have to spend.  In the future, we'd take into account and create family income on the monthly level and not just industry members of household worked in, but account for their job.  Future research is recommended to account for jobs within industries that can be performed from an employee's home.  Doing so would assist in identifying and quantifying how many earners moved from cities to suburbs or rural areas.  Cohorts that relocated during the pandemic to suburbs and working in high-earning occupations present an opportunity for businesses to meet customers demands at different geographic locations.

The below table summarizes our regression of type of industry respondent claimed to work and time period on reported employment. 


| Industry |  % Change
:------    |:-------------------------------------------------     | 
`Arts,Entertainment,andRecreation,andAccommodationandFoodServices`                  | -%05.22 |
`Construction`                                                                      | -%00.95 |
`EducationalServices,andHealthCareandSocialAssistance`                              | %01.59 |
`FinanceandInsurance,andRealEstateandRentalandLeasing`                              | %02.37 |
`Information`                                                                       | %00.02 |
`Manufacturing`                                                                     | %00.81 |
`OtherServices,ExceptPublicAdministration`                                          | -%00.10 |
`Professional,Scientific,andManagement,andAdministrativeandWasteManagementServices` | %00.32 |
`PublicAdministration`                                                              | %03.07 |
`RetailTrade`                                                                       | -%00.66 |
`TransportationandWarehousing,andUtilities`                                         | -%00.13 |
`WholesaleTrade`                                                                    | %01.36 |
`period_typeCovid`                                                                  | -%00.81 |
`period_typePost_Covid`                                                             | -%2.92


Those census survey respondents that work as public administration for state and local governments, finance and real estate, technology and specialized trades are most likely to be employed.  This strata lives predominantly in the suburbs and city centers of major east and west coast states offering jobs. Again our assumption that being employed means money to spend and fill the needs for consumers.  We recommend it a benefit to review industry type and ability for employees to work remotely and how geographic re-location is/has changing is accordance with remote work.  Reviewing debt and savings rates for consumers and accounting for children in household would be beneficial in determining the true amount of income a family has to spend.     

## Load Libraries
```{r, warning = FALSE, message = FALSE}
# Load Relevant Libraries
library(ipumsr)  
library(dplyr)
library(tidyverse)
library(vtable)
library(lubridate)
library(haven)
library(fixest)
library(ggplot2)
library(formattable)
library(ggplot2)
```

## Load and Mutate Data

```{r, warning = FALSE, message = FALSE}
#### READ IN INDUSTRY CODES ####
Industries <- read_csv('indnames.csv') %>%
  rename(IND = ind)
###### Read in State Federal Information Processing Codes. ######  
state_fips <- read_csv('state-geocodes-v2021.csv') %>%
  filter(!is.na(Region)) %>%
  rename(State = Name)
  #select(STATEFIP, State)
###### read in IPUMS demographic data from
dem_ddi <- read_ipums_ddi("cps_00016.xml")
dem_data <- read_ipums_micro(dem_ddi) %>% 
  zap_labels() %>%  
  mutate(
    month_digit= case_when( 
      nchar(str_trim(as.character(MONTH))) ==1 ~ (paste('0',str_trim(as.character(MONTH)), sep = '')),
      nchar(str_trim(as.character(MONTH))) ==2 ~ (str_trim(as.character(MONTH))), 
      TRUE ~ 'WRONG'
    )
    , date_ = as_date(paste(YEAR, month_digit, '01', sep = '-'))
    , period_type = as_factor(case_when(
      date_ < '2020-03-01' ~'Pre_Covid', 
      date_ >= '2020-03-01' & date_ < '2020-04-01' ~'Covid', 
      date_ >= '2020-04-01' ~'Post_Covid'
    ))
  )
#### JOIN FILES INTO ONE ####
DF <- dem_data %>% 
  ## Join in geographic info
  left_join(state_fips, by = 'STATEFIP') %>% # 3.9 million observations
  ## Join in industry categories
  left_join(Industries, by = 'IND')

#### DEMOGRAPHICS DATAFRAME ####
demo_data <- DF %>%
  mutate(EmploymentStatus = case_when(
    EMPSTAT <= 1 | EMPSTAT >= 30 ~  'Other',
    EMPSTAT >= 10 & EMPSTAT < 20 ~ '1',
    EMPSTAT >= 20 & EMPSTAT < 30 ~ '0')) %>%
  mutate(sex = case_when(
    SEX == 1 ~  'male',
    SEX == 2 ~ 'female')) %>%
  mutate(Married = case_when(
    MARST == 1 ~ 'Married',
    MARST != 1 ~ 'Single')) %>%
  mutate(RACE = case_when(
    HISPAN == '0' & RACE == '100' ~ 'White',
    HISPAN == '0' & RACE == '200' ~ 'Black',  
    HISPAN == '0' & RACE == '300' ~ 'AmericanIndian',
    HISPAN == '0' & RACE %in% c('650', '651', '652') ~ 'Asian',
    HISPAN == '0' & !(RACE %in% c('100', '200', '300', '650', '651', '652')) ~ 'MixedRace',
    HISPAN != '0' ~ 'Latino')) %>%
  mutate(AgeGroup = case_when(
    AGE < 20 ~ 'under20',
    AGE >= 20 & AGE < 30 ~ '20to29',
    AGE >= 30 & AGE < 40 ~ '30to39',
    AGE >= 40 & AGE < 50 ~ '40to49',
    AGE >= 50 & AGE < 60 ~ '50to59',
    AGE >= 60 & AGE <= 65 ~ '60to65',
    AGE >= 65 ~ 'RetirementAge')) %>%
  mutate(REGION = case_when(
    REGION == 11 | REGION == 12 ~ 'Northeast',
    REGION == 21 | REGION == 22 ~ 'Midwest',
    REGION == 31 | REGION == 32 |  REGION == 33 ~ 'South',
    REGION == 41 | REGION == 42 ~ 'West',
    REGION == 97 ~ 'Unknown')) %>%
  mutate(METRO = case_when (
    METRO %in% c(0, 4) ~  'Other',
    METRO == 1 ~  'Rural',
    METRO == 2 ~  'CityCenter',
    METRO == 3 ~  'Suburb')) %>%
  mutate(Education = case_when(
    EDUC99 == 00 ~ 'Unknown',
    EDUC99 >= 1 & EDUC99 <=9 ~ 'NoDiploma',
    EDUC99 == 10 ~ 'HighSchoolDiploma',
    EDUC99 == 11 ~ 'SomeCollege',
    EDUC99 >= 12 & EDUC99 <= 14 ~  'AssociateDegree',
    EDUC99 == 15 ~ 'Bachelors',
    EDUC99 == 16 ~ 'Masters',
    EDUC99 > 16 ~ 'PhDProfessional')) %>%
  mutate(LaborForce = case_when(
    LABFORCE == 0  ~ 'Military',
    LABFORCE == 1  ~ '0',
    LABFORCE == 2  ~ '1')) %>%
  mutate(WorkStatus = case_when(
    WKSTAT %in% c(11, 14, 15) ~ 'FullTime',
    WKSTAT %in% c(12, 20, 21, 22, 40, 41) ~ 'PartTime',
    WKSTAT == 13 ~ 'FullTimeNoWork',
    WKSTAT == 42 ~ 'PartTimeNoWork',
    WKSTAT == 50 ~ 'NoWorkSeekingFullTime',
    WKSTAT == 60 ~ 'NoWorkSeekingPartTime',
    WKSTAT == 99 ~ 'MilitaryNoLaborForce',)) %>%
  mutate(FamilyIncome =case_when(                        
    FAMINC <= 600 ~ '<$25k_FPL_Familyof4',
    FAMINC > 600 & FAMINC <= 740 ~ '25k-50K',
    FAMINC > 740 & FAMINC <= 830 ~ '50k-75k',
    FAMINC > 830 & FAMINC <= 841~ '75k-100k',
    FAMINC == 842  ~ '100k - 150k',
    FAMINC == 843 ~ '150k+',
    FAMINC == 999 ~ 'NA')) %>%
  mutate(WHYUNEMP = case_when(
    WHYUNEMP <= 3 ~ 'LostJob',
    WHYUNEMP == 4 ~ 'Quit',
    WHYUNEMP == 5 ~ 'ReEnterLaborForce',
    WHYUNEMP == 6 ~ 'FirstJob',
    WHYUNEMP == 0 ~ 'Military')) %>%
  mutate(employed = ifelse(EMPSTAT >= 10 & EMPSTAT <20, 1, 0), unemployed = ifelse(EMPSTAT >= 20 & EMPSTAT<= 22, 1, 0)) %>%  
  mutate(is_employed = ifelse(EMPSTAT>=10 & EMPSTAT<20, TRUE, FALSE)) %>%  
  mutate(retail_v_other = ifelse(indname=='Retail Trade', 'Retail', 'Other') ## flag if industry == Retail
  )%>% 
  filter(EMPSTAT < 30 & EMPSTAT >= 10) %>% 
  mutate_if(is.character, as.factor) %>%
  unite(HHID, HRHHID, HRHHID2) %>%
  drop_na(indname) %>%  # Dropping N/A's in indname as NIU's in military.
  select(date_, YEAR, MONTH, HHID, EMPSTAT, is_employed, employed, unemployed, CPSIDP
          , HWTFINL, WTFINL, FAMINC, REGION, METRO, AGE, AgeGroup, SEX, sex
          , RACE, Married, IND, WHYUNEMP, AgeGroup, Education, LaborForce
          , EmploymentStatus, WorkStatus, FamilyIncome, indname, retail_v_other, State, period_type) %>%
  rename(final_weight= WTFINL, person_id = CPSIDP, household_weight = HWTFINL)

```


## EDA

```{r, warning = FALSE, message = FALSE}
# distinct people in survey = 500,130
summary(demo_data)
vtable(demo_data, lush = TRUE)
reg_df_vis<- demo_data %>% 
  select(period_type
         , date_
         ## Measure
         , is_employed
         , final_weight
         ## Variables
         , IND, indname, REGION
         ## Fixed
         , AGE, sex, RACE
  )
```



```{r, warning = FALSE, message = FALSE}
## All industries unemployment rate over time
unemp_time <- demo_data %>% 
  filter(indname != 'Military', indname != 'Other') %>%
  group_by(date_) %>%
  summarize(UnemploymentRate = 1- (weighted.mean(is_employed == TRUE, w = final_weight))) %>% 
  ggplot(aes(x = date_, y = UnemploymentRate)) + geom_line()+
  geom_rect(aes(xmin=as_date('2020-03-01'), xmax=as_date('2020-04-01'), ymin=-Inf, ymax=Inf),  linetype = 'dashed',color="light green", fill = 'light green' ,alpha=.01)+
  scale_y_continuous(labels = scales::percent_format(accuracy = .1))+
  labs(x = "Survey Date", y = "Unemployment Rate",
    title ="Unemployment - All Industry")+
  theme(aspect.ratio=3/7)
unemp_time

## All industries v retail unemployment rate over time
unemp_retail_time <- demo_data %>%
  filter(indname != 'Military', indname != 'Other') %>% 
  group_by(date_, retail_v_other) %>%
  summarize(UnemploymentRate = 1- (weighted.mean(is_employed == TRUE, w = final_weight))) %>% 
  ggplot(aes(x = date_, y = UnemploymentRate, color = retail_v_other)) + geom_line()+
  geom_rect(aes(xmin=as_date('2020-03-01'), xmax=as_date('2020-04-01'), ymin=-Inf, ymax=Inf),  linetype = 'dashed',color="light green", fill = 'light green' ,alpha=.01)+
  labs(x = "Survey Date", y = "Unemployment Rate",
    title ="Unemployment (Retail v Other Industries)")+
  scale_y_continuous(labels = scales::percent_format(accuracy = .1))+
  theme(aspect.ratio=3/7)
unemp_retail_time
## limit x window for clarity
unemp_time_skinny <- demo_data %>% 
  filter(date_>=as_date('2020-01-01'),
         date_<= as_date('2020-06-01')) %>%
  filter(indname != 'Military', indname != 'Other') %>%
  group_by(date_, retail_v_other) %>%
  summarize(UnemploymentRate = 1- (weighted.mean(is_employed == TRUE, w = final_weight))) %>% 
  ggplot(aes(x = date_, y = UnemploymentRate, color = retail_v_other)) + geom_line()+
  geom_rect(aes(xmin=as_date('2020-03-01'), xmax=as_date('2020-04-01'), ymin=-Inf, ymax=Inf),  linetype = 'dashed',color="light green", fill = 'light green' ,alpha=.01)+
  labs(x = "Survey Date", y = "Unemployment Rate",
    title ="Unemployment (Retail v Other Industries), 2020")+
  scale_y_continuous(labels = scales::percent_format(accuracy = .1))+
  theme(aspect.ratio=3/7)
unemp_time_skinny
```

```{r, warning = FALSE, message = FALSE}
unemp_industries <- demo_data %>%
  filter(indname != 'Military', indname != 'Other') %>%
  group_by(date_, indname) %>%
  summarize(UnemploymentRate = 1- (weighted.mean(is_employed == TRUE, w = final_weight))) %>% 
  ggplot(aes(x = date_, y = UnemploymentRate, color = indname)) + geom_line()+
  geom_rect(aes(xmin=as_date('2020-03-01'), xmax=as_date('2020-04-01'), ymin=-Inf, ymax=Inf),  linetype = 'dashed',color="light green", fill = 'light green' ,alpha=.01)+ labs(x = "Survey Date", y = "Unemployment Rate",
    title ="Unemployment All Industries")+
  scale_y_continuous(labels = scales::percent_format(accuracy = .1))+
  theme(aspect.ratio=3/7, legend.position = 'bottom')
unemp_industries
```

### Geographical
```{r, warning = FALSE, message = FALSE}
unemp_by_region <- demo_data %>%
  filter(indname != 'Military', indname != 'Other') %>% 
  group_by(date_, REGION) %>%
  summarize(UnemploymentRate = 1- (weighted.mean(is_employed == TRUE, w = final_weight))) %>% 
  ggplot(aes(x = date_, y = UnemploymentRate, color = REGION)) + geom_line()+
  geom_rect(aes(xmin=as_date('2020-03-01'), xmax=as_date('2020-04-01'), ymin=-Inf, ymax=Inf),  linetype = 'dashed',color="light green", fill = 'light green' ,alpha=.01)+
  labs(x = "Survey Date", y = "Unemployment Rate",
    title ="Unemployment by Region All Industries")+
  scale_y_continuous(labels = scales::percent_format(accuracy = .1))+
  theme(aspect.ratio=3/7)
unemp_by_region

unemp_by_metro <- demo_data %>%
  filter(indname != 'Military', indname != 'Other') %>% 
  group_by(date_, METRO) %>%
  summarize(UnemploymentRate = 1- (weighted.mean(is_employed == TRUE, w = final_weight))) %>% 
  ggplot(aes(x = date_, y = UnemploymentRate, color = METRO)) + geom_line()+
  theme(aspect.ratio=3/7)+
  geom_rect(aes(xmin=as_date('2020-03-01'), xmax=as_date('2020-04-01'), ymin=-Inf, ymax=Inf),  linetype = 'dashed',color="light green", fill = 'light green' ,alpha=.01)+
  labs(x = "Survey Date", y = "Unemployment Rate",
    title ="Unemployment by Location : City, Suburban, Rural, Other")+
  scale_y_continuous(labels = scales::percent_format(accuracy = .1))+
  theme(aspect.ratio=3/7)
unemp_by_metro
```
During March 2020, all regions experienced sudden and parallel changes in unemployment rates and those rates, regardless if living in the city, suburbs or rural country side experienced similar unemployment rate changes.  Differences in the decline of unemployment rates from April 2020 through early 2021 are remarkable for higher rates of unemployment in the Northeast and Western United States.  These regions are populated with larger cities, thus it is not surprising to that unemployment in the cities was highest during this period.

### Race
```{r, warning = FALSE, message = FALSE}
#### RACE PRE-COVID EMPLOYMENT ####
unemp_by_race <- demo_data %>%
  filter(indname != 'Military', indname != 'Other') %>% 
  group_by(date_, RACE) %>%
  summarize(UnemploymentRate = 1- (weighted.mean(is_employed == TRUE, w = final_weight))) %>% 
  ggplot(aes(x = date_, y = UnemploymentRate, color = RACE)) + geom_line()+
  theme(aspect.ratio=3/7)+
  geom_rect(aes(xmin=as_date('2020-03-01'), xmax=as_date('2020-04-01'), ymin=-Inf, ymax=Inf),  linetype = 'dashed',color="light green", fill = 'light green' ,alpha=.01)+
  labs(x = "Survey Date", y = "Unemployment Rate",
    title ="Unemployment All Industries")+
  scale_y_continuous(labels = scales::percent_format(accuracy = .1))+
  theme(aspect.ratio=3/7)
unemp_by_race
```
Evaluating all industry types, American Indians, African-Americans and Latino's experienced higher unemployment rates vs Caucasians and people of Asian descent

### Age
```{r, warning = FALSE, message = FALSE}
unemp_by_age_group <- demo_data %>%
  filter(indname != 'Military', indname != 'Other') %>% 
  group_by(date_, AgeGroup) %>%
  summarize(UnemploymentRate = 1- (weighted.mean(is_employed == TRUE, w = final_weight))) %>% 
  ggplot(aes(x = date_, y = UnemploymentRate, color = AgeGroup)) + geom_line()+
  theme(aspect.ratio=3/7)+
  geom_rect(aes(xmin=as_date('2020-03-01'), xmax=as_date('2020-04-01'), ymin=-Inf, ymax=Inf),  linetype = 'dashed',color="light green", fill = 'light green' ,alpha=.01)+
  labs(x = "Survey Date", y = "Unemployment Rate",
    title ="Unemployment All Industries")+
  scale_y_continuous(labels = scales::percent_format(accuracy = .1))+
  theme(aspect.ratio=3/7)
unemp_by_age_group
```
People less than 30 years old experienced higher rates of unemployment vs. other age categories.  The 30-39, 40-49, and 50-59 age groups experienced the lowest rates of unemployment in all industries

# Regressions
```{r, warning = FALSE, message = FALSE}
reg_df <- demo_data %>%
  select(period_type
         , date_
         ## Measure
         , is_employed
         , final_weight
         ## Variables
         , IND, indname, REGION
         ## Fixed
         , AGE, sex, RACE, AgeGroup, Married
  )

reg1 <- feols(is_employed~period_type + indname, se = 'hetero', data = reg_df)
etable(reg1)               
```
We ran regression 1 (reg1) to evaluate while holding industry type constant, the effect of time (pre-covid, covid, and post-covid) on whether or not respondents were employed.  Our assumptions were if one had a job, there a likelihood they would have money to spend, all jobs are created equal irrespective of industry, and even though some industries had workers that changed the location of where they produced, controlling for industry was sufficient to look at simply who had jobs.  The intent was to identify number of jobs in respect to three different time periods and control for industry with the assumption that even though some industries had changes as to where workers completed work.  Jobs within Public Administration and Finance have done well and marketing goods and services to that strata would be pragmatic as they have a job and can spend money.  The military industry should be ignored and is likely an error in data wrangling as the coefficient does not make logical sense. 

-Interpretation reg1 =  

-controlling for industry and relative to agricultural industry, a one-unit change in pre-covid employment is associated with a -5.22% decrease in within the Entertainment and Recreation industry.

-controlling for industry and relative to agricultural industry, a one-unit change in pre-covid employment is associated with a -.95% decrease within the Educational Services industry.

-controlling for industry and relative to agricultural industry, a one-unit change in pre-covid employment is associated with a -1.59% increase within the Educational Services industry.

-controlling for industry and relative to agricultural industry, a one-unit change in pre-covid employment is associated with a 2.37% increase within the Financial & Insurance services industry.

-controlling for industry and relative to agricultural industry, a one-unit change in pre-covid employment is associated with a .02% increase within the Information services industry.

-controlling for industry and relative to agricultural industry, a one-unit change in pre-covid employment is associated with a .81% increase within the Manufacturing industry.

-controlling for industry and relative to agricultural industry, a one-unit change in pre-covid employment is associated with a -95.02% decrease within the Manufacturing industry.  # logically work through this, does it make sense?  Simply filter out?

-controlling for industry and relative to agricultural industry, a one-unit change in pre-covid employment is associated with a -.1% decrease within Other Services Industry.

-controlling for industry and relative to agricultural industry, a one-unit change in pre-covid employment is associated with a .32% increase within Professional, Scientific and Management Industry.

-controlling for industry and relative to agricultural industry, a one-unit change in pre-covid employment is associated with a 3.7% increase within Public Administration Industry.

-controlling for industry and relative to agricultural industry, a one-unit change in pre-covid employment is associated with a -.66% decrease within Retail Trade Industry.

-controlling for industry and relative to agricultural industry, a one-unit change in pre-covid employment is associated with a -.13% decrease within Transportation & warehousing Industry.

-controlling for industry and relative to agricultural industry, a one-unit change in pre-covid employment is associated with a 1.36% increase within Wholesale Trade Industry.


```{r}
reg2 <- feols(is_employed~indname + period_type, se = 'hetero', data = reg_df)
etable(reg2)
```

We have assumed that jobs equal money to solicit good and services to customers and that all jobs, regardless of industry are the same.  Interpretations for some industries of interest in  reg2 (please note:summary table above in findings).  Note that values are relative to agriculture and period_type is relative to pre-covid.

-controlling for period_type and relative to agricultural industry, a one-unit change in industry employment is associated with a -5.22% decrease within the Art, Entertainment and Recreation industry.
 
-controlling for period_type and relative to agricultural industry, a one-unit change in industry employment is associated with a -.66% decrease within the Retail Trade industry.
 
-controlling for period_type and relative to agricultural industry, a one-unit change in industry employment is associated with a -.13% decrease within the Transportation & warehousing industry.

It is a bit presumptuous that a percentage increase or decrease in one industry where workers report being employed in a traditionally highly paid industry is weighted similarly to a low-skilled job paying minimum wage.  The model could be improved by evaluating occupational code names from Census Bureau, evaluating median FAMINC at the annual-level of respondents that worked certain occupations, and then assigning weights to the jobs unemployment figures.  Cross-referencing industry by geography could aid in determining where high-paying industries are clustered, and assuming those people working within those industries have money to spend, it would be prudent to market goods and services to that target market.
