---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

  
```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = TRUE)

    knitr::opts_knit$set(root.dir = normalizePath("/Users/khigashi/Documents/OMSBA/OMSBA-5300-DTC-Erik-collab/data"))
```

```{r}

library(ipumsr)  
library(dplyr)
library(tidyverse)
library(jtools)
library(vtable)
library(lubridate)
library(haven)
library(fixest)
library(ggplot2)
library(formattable)
library(echarts4r)
library(ggplot2)
```

# Read in Data
```{r}

#### READ IN INDUSTRY CODES ####
Industries <- read_csv('indnames.csv') %>%
  rename(IND = ind)

###### Read in State Federal Information Processing Codes. ######  
state_fips <- read_csv('state-geocodes-v2021.csv') %>%
  filter(!is.na(Region)) %>%
  rename(State = Name) %>%
  select(STATEFIP, State)

###### read in IPUMS demographic data from
dem_ddi <- read_ipums_ddi("cps_00016.xml")
dem_data <- read_ipums_micro(dem_ddi) %>% 
  zap_labels() %>%  
  mutate(
    month_digit= case_when( 
      nchar(str_trim(as.character(MONTH))) ==1 ~ (paste('0',str_trim(as.character(MONTH)), sep = '')),
      nchar(str_trim(as.character(MONTH))) ==2 ~ (str_trim(as.character(MONTH))), 
      TRUE ~ 'WRONG'
    )
    , date_ = as_date(paste(YEAR, month_digit, '01', sep = '-'))
    , period_type = as_factor(case_when(
      date_ < '2020-03-01' ~'Pre_Covid', 
      date_ >= '2020-03-01' & date_ < '2020-04-01' ~'Covid', 
      date_ >= '2020-04-01' ~'Post_Covid'
    ))
  )

#### JOIN FILES INTO ONE ####
DF <- dem_data %>% 
  ## Join in geographic info
  left_join(state_fips, by = 'STATEFIP') %>% # 3.9 million observations
  ## Join in industry categories
  left_join(Industries, by = 'IND')

#### DEMOGRAPHICS DATAFRAME ####
demo_data <- DF %>%
  mutate(EmploymentStatus = case_when(
    EMPSTAT <= 1 | EMPSTAT >= 30 ~  'Other',
    EMPSTAT >= 10 & EMPSTAT < 20 ~ '1',
    EMPSTAT >= 20 & EMPSTAT < 30 ~ '0')) %>%
  mutate(sex = case_when(
    SEX == 1 ~  'male',
    SEX == 2 ~ 'female')) %>%
  mutate(MARST = case_when(
    MARST == 1 ~ 'Married',
    MARST != 1 ~ 'Single')) %>%
  mutate(RACE = case_when(
    HISPAN == '0' & RACE == '100' ~ 'White',
    HISPAN == '0' & RACE == '200' ~ 'Black',  
    HISPAN == '0' & RACE == '300' ~ 'AmericanIndian',
    HISPAN == '0' & RACE %in% c('650', '651', '652') ~ 'Asian',
    HISPAN == '0' & !(RACE %in% c('100', '200', '300', '650', '651', '652')) ~ 'MixedRace',
    HISPAN != '0' ~ 'Latino')) %>%
  mutate(AgeGroup = case_when(
    AGE < 20 ~ 'under20',
    AGE >= 20 & AGE < 30 ~ '20to29',
    AGE >= 30 & AGE < 40 ~ '30to39',
    AGE >= 40 & AGE < 50 ~ '40to49',
    AGE >= 50 & AGE < 60 ~ '50to59',
    AGE >= 60 & AGE <= 65 ~ '60to65',
    AGE >= 65 ~ 'RetirementAge')) %>%
  mutate(REGION = case_when(
    REGION == 11 | REGION == 12 ~ 'Northeast',
    REGION == 21 | REGION == 22 ~ 'Midwest',
    REGION == 31 | REGION == 32 |  REGION == 33 ~ 'South',
    REGION == 41 | REGION == 42 ~ 'West',
    REGION == 97 ~ 'Unknown')) %>%
  mutate(METRO = case_when (
    METRO %in% c(0, 4) ~  'Other',
    METRO == 1 ~  'Rural',
    METRO == 2 ~  'CityCenter',
    METRO == 3 ~  'Suburb')) %>%
  mutate(Education = case_when(
    EDUC99 == 00 ~ 'Unknown',
    EDUC99 >= 1 & EDUC99 <=9 ~ 'NoDiploma',
    EDUC99 == 10 ~ 'HighSchoolDiploma',
    EDUC99 == 11 ~ 'SomeCollege',
    EDUC99 >= 12 & EDUC99 <= 14 ~  'AssociateDegree',
    EDUC99 == 15 ~ 'Bachelors',
    EDUC99 == 16 ~ 'Masters',
    EDUC99 > 16 ~ 'PhDProfessional')) %>%
  mutate(LaborForce = case_when(
    LABFORCE == 0  ~ 'Military',
    LABFORCE == 1  ~ '0',
    LABFORCE == 2  ~ '1')) %>%
  mutate(WorkStatus = case_when(
    WKSTAT %in% c(11, 14, 15) ~ 'FullTime',
    WKSTAT %in% c(12, 20, 21, 22, 40, 41) ~ 'PartTime',
    WKSTAT == 13 ~ 'FullTimeNoWork',
    WKSTAT == 42 ~ 'PartTimeNoWork',
    WKSTAT == 50 ~ 'NoWorkSeekingFullTime',
    WKSTAT == 60 ~ 'NoWorkSeekingPartTime',
    WKSTAT == 99 ~ 'MilitaryNoLaborForce',)) %>%
  mutate(FamilyIncome =case_when(                        
    FAMINC <= 600 ~ '<$25k_FPL_Familyof4',
    FAMINC > 600 & FAMINC <= 740 ~ '25k-50K',
    FAMINC > 740 & FAMINC <= 830 ~ '50k-75k',
    FAMINC > 830 & FAMINC <= 841~ '75k-100k',
    FAMINC == 842  ~ '100k - 150k',
    FAMINC == 843 ~ '150k+',
    FAMINC == 999 ~ 'NA')) %>%
  mutate(WHYUNEMP = case_when(
    WHYUNEMP <= 3 ~ 'LostJob',
    WHYUNEMP == 4 ~ 'Quit',
    WHYUNEMP == 5 ~ 'ReEnterLaborForce',
    WHYUNEMP == 6 ~ 'FirstJob',
    WHYUNEMP == 0 ~ 'Military')) %>%
  mutate(employed = ifelse(EMPSTAT <20, 1, 0), unemployed = ifelse(EMPSTAT > 20, 1, 0)) %>% 
  mutate(is_employed = ifelse(EMPSTAT<20 & EMPSTAT>10, TRUE, FALSE)
  ) %>% 
  filter(EMPSTAT < 30 & EMPSTAT >= 10) %>% 
  mutate_if(is.character, as.factor) %>%
  unite(HHID, HRHHID, HRHHID2) %>%     #When combined with HRHHID2, HRHHID can uniquely identify households within basic monthly samples-IPUMS
  # mutate(HHID = as.numeric(HHID)) %>%
  drop_na(indname) %>% # Dropping N/A's in indname as NIU's in military.
  select(date_, YEAR, MONTH, HHID, PERNUM, EMPSTAT, is_employed, CPSID, CPSIDP
         , HWTFINL, WTFINL, FAMINC, REGION, METRO, AGE, SEX, sex
         , RACE, MARST, IND, WHYUNEMP, AgeGroup, Education, LaborForce
         , EmploymentStatus, WorkStatus, FamilyIncome, indname, State, period_type)
```



# Regressions 
Notes: 

Our hypothesis: Covid disproportionately impacted people of color. Has money == Has Job == Spends money == Happy Retail

Null hypothesis: Covid did NOT impact employment industries regionally.

Treatment = during/post Covid
Control == precovid

Fixed: Age, Sex, Race, == Human demographics (age since 
      time period is only 1-2 years)

Variables:
  Region
  Industry
  EmploymentStatus
Outcome: employmentStatus

employment = B0 +IndustryB1 + RegionB2 + e

```{r}
head(demo_data)

# 
# EMP_INDUSTRY <- demo_data %>%
#   group_by(date_, indname, period_type) %>%
#   summarize(UmploymentRateIndustry = 1- (weighted.mean(employed == 1, w = WTFINL))) %>%
#   filter(indname != 'Military', indname != 'Other')



reg_df <- demo_data %>% 
  select(period_type
         , date_
         ## Measure
         , is_employed
         , WTFINL
         ## Variables
         , IND, indname, REGION
         ## Fixed
         , AGE, sex, RACE
  )
summary(reg_df)

reg1 <- reg_df %>% 
  feols(is_employed~period_type, se = 'hetero')

reg2 <- reg_df %>% 
  feols(is_employed~period_type+ RACE, se = 'hetero')
reg3 <- reg_df %>% 
  feols(is_employed~period_type | RACE)
etable(reg2, reg3)
reg4 <- reg_df %>% 
  feols(is_employed~period_type | AGE)
reg2

etable(reg1, reg2, reg3, reg4)

``` 



<!-- # Explatory Data Analysis -->
<!-- ```{r} -->
<!-- # distinct people in survey = 500,130 -->
<!-- distinct_people <- demo_data %>% -->
<!--   distinct(CPSIDP) -->
<!-- summary(demo_data %>% -->
<!--           select(-YEAR, -MONTH, -date_, -PERNUM, -PERNUM, -HWTFINL, -WTFINL, -IND, - EMPSTAT,  -->
<!--                  -CPSID, -CPSIDP)) -->

<!-- # distinct people in survey = 500,130 -->
<!-- distinct_people <- demo_data %>% -->
<!--   distinct(CPSIDP) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- #### EMPLOYMENT & UNEMPLOYMENT ######## -->
<!-- ##### INDUSTRY-PRE-COVID -ALL ##### -->

<!-- EMP_INDUSTRY <- demo_data %>% -->
<!--   filter(EMPSTAT < 30 & EMPSTAT >= 10) %>%  -->
<!--   mutate(employed = ifelse(EMPSTAT <20, 1, 0) -->
<!--          , unemployed = ifelse(EMPSTAT > 20, 1, 0)) %>% -->
<!--   group_by(date_, indname, period_type) %>% -->
<!--   summarize(UmploymentRateIndustry = 1- (weighted.mean(employed == 1, w = WTFINL))) %>% -->
<!--   filter(indname != 'Military', indname != 'Other') -->

<!-- ``` -->

<!-- ```{r} -->
<!-- ##### METRO-PRE-COVID ##### -->

<!-- EMP_Metro <- demo_data %>% -->
<!--   filter(EMPSTAT < 30 & EMPSTAT >= 10) %>%  -->
<!--   mutate(employed = ifelse(EMPSTAT <20, 1, 0) -->
<!--          , unemployed = ifelse(EMPSTAT > 20, 1, 0)) %>% -->
<!--   group_by(date_, period_type, METRO) %>% -->
<!--   summarize(UnemploymentRate = 1- (weighted.mean(employed == 1, w = WTFINL))) %>%  -->
<!--   ungroup() -->


<!-- ``` -->


<!-- ```{r} -->
<!-- ##### Construct chartS for Pre-covid unemployment by METRO, SUBURBAN, RURAL ##### -->
<!-- EMPC <- EMP_Metro %>% -->
<!--   pivot_wider(1, names_from = "METRO", values_from = "UnemploymentRate") -->

<!-- EMPC %>% -->
<!--   ungroup() %>% # Sometime ungroup is needed after doing group_by() ? -->
<!--   e_charts(date_) %>% -->
<!--   e_line(CityCenter, color="#333") %>% -->
<!--   e_line(Other, color="blue") %>% -->
<!--   e_line(Rural, color="green") %>% -->
<!--   e_line(Suburb, color="orange") %>% -->
<!--   e_tooltip(trigger = "axis") -->

<!-- EMPC_avg <- EMPC %>% -->
<!--   group_by() %>% -->
<!--   summarise( -->
<!--     CityCenter = mean(CityCenter), -->
<!--     Other = mean(Other), -->
<!--     Rural = mean(Rural), -->
<!--     Suburb = mean(Suburb)) %>% -->
<!--   ungroup() -->

<!-- #create a dataframe with values to chart -->
<!-- EMPC_avg_bar <- data.frame( -->
<!--   names = c("CityCenter", "Other", "Rural", "Suburb"), -->
<!--   values = c(0.0379, 0.0353, 0.0383, 0.0301)) -->

<!-- ##### Employment average by CityCenter, Rural, Suburb ##### -->
<!-- EMPC_avg_bar %>% -->
<!--   e_charts(names) %>% -->
<!--   e_bar(values) -->

<!-- ##### METRO-COVID ##### -->
<!-- EMP_Metro_Covid <- demo_data %>% -->
<!--   filter(EMPSTAT < 30 & EMPSTAT >= 10) %>%  -->
<!--   mutate(employed = ifelse(EMPSTAT <20, 1, 0), unemployed = ifelse(EMPSTAT > 20, 1, 0)) %>% -->
<!--   group_by(METRO, period_type) %>%  -->
<!--   summarize(UnemploymentRate = 1- (weighted.mean(employed == 1, w = WTFINL))) -->
<!-- summary(EMP_Metro_Covid) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- ##### Construct chartS for COVID unemployment by METRO, SUBURBAN, RURAL ##### -->
<!-- EMPMC <- EMP_Metro_Covid %>% -->
<!--   ungroup() -->

<!-- EMPMC %>% -->
<!--   e_charts(METRO) %>% -->
<!--   e_bar(UnemploymentRate) %>% -->
<!--   e_tooltip() -->

<!-- ##### Construct chartS for POST-COVID unemployment by METRO, SUBURBAN, RURAL ##### -->

<!-- EMPMPC <- EMP_Metro_PostCovid %>%  -->
<!--   ungroup() -->

<!-- EMPMPC %>% -->
<!--   e_charts(METRO) %>% -->
<!--   e_bar(UnemploymentRate) %>% -->
<!--   e_tooltip() -->
<!-- ``` -->
<!-- ``` {r} -->
<!-- ################# REGION ##### -->
<!-- ##### REGION-PRE-COVID #### -->

<!-- EMP_REGION_PreCovid <- demo_data %>% -->
<!--   filter(EMPSTAT < 30 & EMPSTAT >= 10) %>%  -->
<!--   mutate(employed = ifelse(EMPSTAT <20, 1, 0), unemployed = ifelse(EMPSTAT > 20, 1, 0)) %>% -->
<!--   group_by(date_, period_type, REGION) %>%  -->
<!--   summarize(UnemploymentRate = 1- (weighted.mean(employed == 1, w = WTFINL)))%>%  # ungroup -->
<!--   ungroup() %>%  #need to pivot wider -->
<!--   pivot_wider(1, names_from = "REGION", values_from = "UnemploymentRate") -->

<!-- ``` -->

<!-- ```{r} -->
<!-- ###### Construct chartS for PRE-COVID unemployment by REGION #### -->


<!-- EMP_REGION_PreCovid <- EMP_REGION_PreCovid %>%  #need to pivot wider -->
<!--   pivot_wider(1, names_from = "REGION", values_from = "UnemploymentRate") -->

<!-- # unemployment by region PRE-COVID line -->
<!-- EMP_REGION_PreCovid %>% -->
<!--   ungroup() %>% # ungroup is needed after doing group_by() -->
<!--   e_charts(date_) %>% -->
<!--   e_line(Midwest) %>% -->
<!--   e_line(Northeast) %>% -->
<!--   e_line(South) %>% -->
<!--   e_line(West, color="orange") %>% -->
<!--   e_tooltip(trigger = "axis") -->

<!-- # mean unemployment pre-covid -->
<!-- EMPRPC <- EMP_REGION_PreCovid %>% -->
<!--   group_by() %>% -->
<!--   summarise( -->
<!--     Midwest = mean(Midwest), -->
<!--     Northeast = mean(Northeast), -->
<!--     South = mean(South), -->
<!--     West = mean(West)) -->

<!-- #create dataframe for above values -->
<!-- EMP_REGION_PreCovid1 <- data.frame( -->
<!--   REGION = c("Midwest", "Northeast", "South", "West"), -->
<!--   UnemploymentRate = c(0.0339, 0.0344, 0.0324, 0.0362) -->
<!-- ) -->

<!-- EMP_REGION_PreCovid1 %>% -->
<!--   ungroup() %>% -->
<!--   e_chart(REGION) %>% -->
<!--   e_bar(UnemploymentRate) -->


<!-- ``` -->

<!-- ```{r} -->
<!-- ##### REGION-COVID #### -->
<!-- EMP_REGION_Covid <- demo_data %>% -->
<!--   filter(EMPSTAT < 30 & EMPSTAT >= 10) %>%  -->
<!--   mutate(employed = ifelse(EMPSTAT <20, 1, 0), unemployed = ifelse(EMPSTAT > 20, 1, 0)) %>% -->
<!--   group_by(date_, period_type, REGION) %>%  -->
<!--   summarize(UnemploymentRate = 1- (weighted.mean(employed == 1, w = WTFINL))) -->

<!-- ###### construct chart of unemployment by REGION for COVID #### -->

<!-- EMP_REGION_Covid %>% -->
<!--   ungroup() %>% -->
<!--   e_chart(REGION) %>% -->
<!--   e_bar(UnemploymentRate) -->
<!-- ``` -->

<!-- ```{r} -->


<!-- ########### RACE ################################################################################ -->
<!-- #### RACE PRE-COVID EMPLOYMENT #### -->
<!-- EMP_RACE_Precovid <- demo_data %>% -->
<!--   filter(EMPSTAT < 30 & EMPSTAT >= 10) %>%  -->
<!--   mutate(employed = ifelse(EMPSTAT <20, 1, 0), unemployed = ifelse(EMPSTAT > 20, 1, 0)) %>% -->
<!--   group_by(indname, date_, period_type, RACE) %>%  -->
<!--   summarize(UnemploymentRate = 1- (weighted.mean(employed == 1, w = WTFINL))) -->
<!-- head(EMP_REGION_PreCovid) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- #### EMP AGGREGATE (SEX/AGE/GROUP) #### -->
<!-- ################ PRE-COVID AGG ################# -->
<!-- ##### race ##### -->
<!-- EMP_AGG_Precovid_race <- demo_data %>% -->
<!--   mutate(employed = ifelse(EMPSTAT <20, 1, 0), unemployed = ifelse(EMPSTAT > 20, 1, 0)) %>% -->
<!--   group_by(date_, period_type, RACE) %>%  -->
<!--   summarize(UnemploymentRate = 1- (weighted.mean(employed == 1, w = WTFINL))) -->

<!-- ##### sex/race ##### -->
<!-- EMP_AGG_Precovid_sex_race <- demo_data %>% -->
<!--   filter(EMPSTAT < 30 & EMPSTAT >= 10) %>%  -->
<!--   mutate(employed = ifelse(EMPSTAT <20, 1, 0), unemployed = ifelse(EMPSTAT > 20, 1, 0)) %>% -->
<!--   group_by(date_, period_type, sex, RACE) %>%  -->
<!--   summarize(UnemploymentRate = 1- (weighted.mean(employed == 1, w = WTFINL))) -->
<!-- head(EMP_AGG_Precovid_sex_race) -->
<!-- ##### sex/AgeGroup ##### -->
<!-- EMP_AGG_Precovid_sex_age <- demo_data %>% -->
<!--   filter(EMPSTAT < 30 & EMPSTAT >= 10) %>%  -->
<!--   mutate(employed = ifelse(EMPSTAT <20, 1, 0), unemployed = ifelse(EMPSTAT > 20, 1, 0)) %>% -->
<!--   group_by(date_, period_type, sex, AgeGroup) %>%  -->
<!--   summarize(UnemploymentRate = 1- (weighted.mean(employed == 1, w = WTFINL))) -->

<!-- ##### race/AgeGroup ##### -->
<!-- EMP_AGG_Precovid_race_age <- demo_data %>% -->
<!--   filter(EMPSTAT < 30 & EMPSTAT >= 10) %>%  -->
<!--   mutate(employed = ifelse(EMPSTAT <20, 1, 0), unemployed = ifelse(EMPSTAT > 20, 1, 0)) %>% -->
<!--   group_by(date_, period_type, RACE, AgeGroup) %>%  -->
<!--   summarize(UnemploymentRate = 1- (weighted.mean(employed == 1, w = WTFINL))) -->

<!-- ``` -->

