---
title: 1. How has COVID affected the health of the retail industry, as measured by
  employment?
author: "Taylor Mack"
date: "8/27/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Q2. How has retail fared relative to other industries?

The below analysis is focused on how the retail industry fared against other industries between 2019-01 to 2022-12. The key assumption is that employment level is the determining factor on whether a industry is successful or not. 

[IPMUS employment data](https://cps.ipums.org/cps/index.shtml) was used in this analysis, with sample dates ranging from January 2019 to December of 2021. The following variables were used:

| Variable |  Definition
|:------    |:-------------------------------------------------     | 
| `year` | survey year |
| `serial` | household serial number |
| `month` | month |
| `hwtfinal` | Household weight, Basic Monthly |            |
| `cpisid`  | CPSID, household record |
| `asecflag` | Flag for ASEC |
| `pernum` | Person number in sample unit  |
| `wtfinl` | Final Basic Weight |
| `cpsidp` | CPSID, person record  |
| `empstat` | Employment status |
| `ind` | Industry Identifier |
| `indname` | Industry name |
| `industry_cat` | Whether the industry is considered to be Retail or Non-Retail |
| `date` | The date of the survey |
| `covid_active` | A binary flag indicating covid was active
                   0 : date < '2020-03' 
                   1 : date >= '2020-03'|
| `emp_count` | The total count of employment |
| `avg_employment` | The average employment amount per month |
| `employment_change` | The amount of employment change from the previous period |
| `percent_change` | The percent change in employment from the previous period  |


```{r include=FALSE}
library(tidyverse)
library(jtools)
library(vtable)
library(haven)
library(lubridate)
library(estimatr)
library(fixest)
library(dplyr)
```

**Raw Data**:
```{r message=FALSE, warning=FALSE}
industry_names <- read.csv('./indnames.csv') %>%
  mutate(industry_cat = case_when(
    indname == 'Retail Trade' ~ 'Retail',
    indname != 'Retail Trade' ~ 'Non-Retail'))
data <- read_dta('./cps_00004.dta.gz')
```

##Creating Q2 Dataframes
```{r}
# joining the initial data read with the industry_name dataframe
df <- inner_join(data,industry_names,by= "ind") %>%
  mutate(cpsid = format(cpsid,scientific= FALSE)) %>%
  mutate(cpsidp = format(cpsidp,scientific= FALSE)) %>%
  mutate(date = as_date(paste(year, month, '01', sep = '-'))) %>%
  mutate(covid_active = case_when(date >= '2020-03-01' ~ 1,
                                  date < '2020-03-01' ~ 0))
#filtering for employed and cpsidp's that are not zero
filtered_df <- df %>% filter(cpsidp > 0) %>%
  filter(empstat <20 & empstat >= 10) %>% #20 and above indicates they are unemployed/no looking
  group_by(date,indname) %>%
  mutate(emp_count = n())



```
##Creating Q2 Dataframes to explore/visualize data
###Creating a dataframe that will be used to graph the dataset to compare retail vs non-retail
```{r}

industry_cat_df <- filtered_df %>%
               ungroup() %>%
               select(c('industry_cat','year','date','covid_active','emp_count')) %>%
               distinct() %>%
               group_by(industry_cat,date) %>%
               mutate(count = sum(emp_count))%>%
               select(-c(5))%>%
               distinct() %>%
               ungroup()%>%
               group_by(industry_cat)%>%
               mutate(percent_change = ((count-lag(count))/lag(count)))%>%
               mutate(percent_change = case_when(
                 percent_change = is.na(percent_change) ~ 0,
                 percent_change == is.na(percent_change) ~ 0,
                 percent_change != is.na(percent_change) ~ percent_change))
#Creating a dataframe that will be used to graph the dataset to compare retail vs non-retail
industry_df <- filtered_df %>%
  ungroup() %>%
  select(c('indname','year','date','covid_active','emp_count')) %>%
  distinct() %>%
  group_by(indname,date) %>%
  mutate(count = sum(emp_count))%>%
  select(-c(5))%>%
  distinct() %>%
  ungroup()%>%
  group_by(indname)%>%
  mutate(percent_change = ((count-lag(count))/lag(count)))%>%
  mutate(percent_change = case_when(
    percent_change = is.na(percent_change) ~ 0,
    percent_change == is.na(percent_change) ~ 0,
    percent_change != is.na(percent_change) ~ percent_change))
```

#Plotting the Datasource

The first graph is displaying the employment change percentage per month by the two industry categories, retail and non-retail

The second graph is displaying the employment change percentage per month by each industry

```{r}
industry_cat_df %>%
  ggplot(aes(x=date,y=percent_change,group=industry_cat,color=industry_cat)) +
  scale_y_continuous(labels=scales::percent_format(accuracy=.1)) +
  geom_point() + geom_line() + theme(aspect.ratio=6/7) +
  labs(x="Date",y="Employment Percentage Change",color="Industry Category",
       title= "Employment Percentage Change by Industry Category")

industry_df %>%
  ggplot(aes(x=date,y=percent_change,group=indname,color=indname)) +
  scale_y_continuous(labels=scales::percent_format(accuracy=.1)) +
  geom_point() + geom_line() + theme(aspect.ratio=1) +
  labs(x="Date",y="Employment Percentage Change",color="Industry Name",
       title= "Employment Percentage Change by Industry")
```
##Creating a regression df for question 2
```{r}
q2_reg_df <- filtered_df %>%
  ungroup()%>%
  group_by(date,industry_cat)%>%
  mutate(avg_employment = mean(emp_count))%>%
  select(c('industry_cat','year','date','covid_active','avg_employment'))%>%
  distinct()%>%
  ungroup()%>%
  group_by(industry_cat)%>%
  mutate(employment_change = ((avg_employment-lag(avg_employment))))%>%
  drop_na()
```

##Q2 Regressions
```{r}
reg1 <- feols(emp_count~indname,se= 'hetero',data= filtered_df)
reg1.1 <- feols(emp_count~indname*covid_active,se='hetero',data= filtered_df)
etable(reg1,reg1.1)

reg2 <- feols(emp_count~industry_cat,se='hetero',data= filtered_df)
reg2.1 <- feols(emp_count~industry_cat*covid_active,se='hetero',data= filtered_df)
etable(reg2,reg2.1)

reg3 <- feols(employment_change~industry_cat,se= 'hetero',data= q2_reg_df)
reg3.1 <- feols(employment_change~industry_cat*covid_active,se= 'hetero',data= q2_reg_df)
etable(reg3,reg3.1)
```







